---
title: "Stone_A_2"
author: "Alexandra Stone"
date: "September 9, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(readr)
library(gapminder)
surveys <- read.csv("surveys.csv", header = T, sep = ",")
```

#Assignment 2

##1.	Write R code to extract the survey observations for the first three months of 1990 using the filter() function. (5 points)

```{r}
filter(surveys, month %in% c(1, 2, 3), year == 1990)
```

##2.	Sort the 1990 winter surveys data by descending order of record ID, then by ascending order of weight.  (10 points)

```{r}
#arrange in descending order of record id
surveys %>%
  filter(month %in% c(12, 1, 2), year == 1990) %>%
  arrange(desc(record_id))

#arrange in ascending order of weight
surveys %>%
  filter(month %in% c(12, 1, 2), year == 1990) %>%
  arrange(weight)
```

##3.	Write code that returns the record_id, sex and weight of all surveyed individuals of Reithrodontomys montanus (RO), (10 points)

```{r}
surveys %>%
  
  #filter only RO individuals
  filter(species_id %in% "RO") %>%
  
  #return record_id, sex and weight
  select(record_id, sex, weight)
```

##4.	Write code that returns the average weight and hindfoot length of Dipodomys merriami (DM) individuals observed in each month (irrespective of the year). Make sure to exclude NA values. (10 points)

```{r}
surveys %>%
  
  #filter for DM individuals and remove NAs
  filter(species_id %in% "DM",!is.na(weight),!is.na(hindfoot_length)) %>%
  group_by(month) %>%
  
  #return avg weight and hindfoot length
  summarise(weight_average = mean(weight),
  hindfoot_average = mean(hindfoot_length))
```

##5.	Write  code that determines the number of individuals by species observed in the winter of 1990. (15 points)

```{r}
surveys %>%
  
  #filter winter months dec, jan, feb
  filter(month %in% c(12, 1, 2),
  year == 1990, !is.na(species_id)) %>%
  group_by(species_id) %>%
  
  #use n = n() with summarise to count oberservations
  summarise(n = n())
```

##6.	Create a dataframe named gapminder_df and mutate it to contain a column that contains the gross domestic product for each row in the data frame. (5 points)

```{r}
#create data set from gapminder
gapminder_df <- as.data.frame(gapminder)

#gross GDP is population * gdpPercap 
gapminder_df <- mutate(gapminder_df, 
  gdpGross =  gdpPercap * pop)

gapminder_df
```

##7.	Calculate the Mean GDP for Cambodia for the years within the dataset. (15 points)

```{r}
gapminder_df %>%
  
  #filter so only looking at Cambodia
  filter(country == "Cambodia") %>%
  
  #use mean() to calculate mean GDP
  summarise(mean(gdpGross))
```

##8.	Find the year with the maximum life expectancy for countries in Asia and arrange them in descending order by year,  The result should contain the country's name, the year and the life expectancy.  (15 points)

```{r}
gapminder_df %>%
  
  #only want countries in Asia
  filter(continent == "Asia") %>%
  select(country, year, lifeExp) %>%
  group_by(year) %>%
  
  #use top_n to get max
  top_n(1, wt = lifeExp) %>%
  
  #arrange in descending order by year
  arrange(desc(year))
```

##9.	Count the number of observations per continent. (5 points)

```{r}
gapminder_df %>%
  group_by(continent) %>%
  
  #use n = n() with summarise to count oberservations
  summarise(n = n())
```

##10.	 Compute the average and median life expectancy and GDP per capita by continent for the years 1952 and 2007.  Should we be optimistic given the results? (10 points)

```{r}
#compute for 1952
gapminder_df %>%
  filter(year == 1952) %>%
  group_by(continent) %>%
  summarise(median(lifeExp),
            mean(lifeExp),
            median(gdpPercap),
            mean(gdpPercap))

#compute for 2007
gapminder_df %>%
  filter(year == 2007) %>%
  group_by(continent) %>%
  summarise(median(lifeExp),
            mean(lifeExp),
            median(gdpPercap),
            mean(gdpPercap))

```
We could be optimistic given the results.  The median life expectancy, mean life expectancy, median gdp per capita and mean gdp per capita all increased in 2007.