---
title: "Stone_A_7B.Rmd"
author: "Alexandra Stone"
date: "October 21, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#set working directory to desktop 
#setwd("c:/Users/user/Desktop")

#load necessary libraries
library(RCurl)
library(ROAuth)
library(streamR)
library(twitteR)

#Download the certificate needed for authentication. This creates a certificate file on the desktop
download.file(url="http://curl.haxx.se/ca/cacert.pem",
destfile="cacert.pem")

#Create a file to collect all the Twitter JSON data received from the API call
outFile <- "tweets_sample.json"

#Set the configuration details to authorize your application to access Twitter data
requestURL <- "https://api.twitter.com/oauth/request_token"
accessURL <- "https://api.twitter.com/oauth/access_token"
authURL <- "https://api.twitter.com/oauth/authorize"
consumerKey <- "CMYIh4nm4eLrf8lFVAgQQVwug"
consumerSecret <- "xyPUFiPVXro123SnXzzKLxmAPxjRGeh78Lgw4H3QRZSSSug5EM"
accessToken <- "1186306587872223233-zgVTQ7dumnE5lBe0kiEyXsqn8vAu9B"
accessTokenSecret<- "jA4lfIRKEMYQWVjzDSBIdVputKGShR4tVmXg74Vtu3e0s"

#Authenticate user via OAuth handshake and save the OAuth certificate to the local disk for future connections. OAuth is an authentication protocol that enables a third-party application to obtain limited access to an HTTP service without sharing passwords.
my_oauth <- OAuthFactory$new( consumerKey=consumerKey,
                              consumerSecret=consumerSecret,
                              requestURL=requestURL,
                              accessURL=accessURL,
                              authURL=authURL)
                             
#my_oauth$handshake(cainfo="cacert.pem")
#Error in function (type, msg, asError = TRUE) : Unknown SSL protocol error in connection to api.twitter.com:443
```

First suggestion: Reinstall packages
I reinstalled the 4 listed packages and reran the code.  It still produced the same error.

Second suggestion: Check to make sure the keys are correct
I double checked the keys and recopy and pasted them into my code.  I reran the code, and it produced the same error.

Third suggestion: Update R.
I updates R using installr.  My RStudio was not the latest version.  I then reinstalled the packages.  I reran the code, and it still produced the same error.

Fourth suggestion: Regenerate the keys
I regenerated the keys, copy and pasted them into the code and ran the code.  It produced the same error.

Fifth suggestion: Check if I am able to read data from another URL
I used the following code to see if I am able to access another URL:

iris <- read.csv(fileURL, header = FALSE, sep = ",", quote = "\"'")
names(iris) <- c("sepal_length", "sepal_width", "petal_length", "petal_width", "class")
head(iris)

I was able to see read data from the website, which told me that I am able to connect to a URL and the issue I am having isn't due to me not being able to connect to every URL.

Sixth suggestion: Set SSL certs globally
I ran:

options(RCurlOptions = list(cainfo = system.file("CurlSSL", "cacert.pem", package = "RCurl")))

The same error was produced.

Seventh suggestion: Update pem file
I used the following code:

system( paste0( "curl http://curl.haxx.se/ca/cacert.pem -o " , tempdir() , "/cacert.pem" ) )
my_oauth$handshake( cainfo = paste0( tempdir() , "/cacert.pem" ) )

This did not work, and when looking on stackoverflow it appears it also didn't work with others who used download.file() earlier in their code.

Eight suggestion: Try running your code outside of RStudio
I just opened  up an R console instead of using RStudio and reran my code.  The same error was produced.  