---
title: "Stone_A_5"
author: "Alexandra Stone"
date: "September 29, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Assignment 5
```{r}
#load packages
library(tidyverse)
library(stringr)
library(readr)
library(readxl)
library(dplyr)
library(forcats)
library(chron)
library(lubridate)

#load datasets
na.vals <- c("", "NA", "n/a", "N/A", "none")
fmarkets <- read_csv("C:/Users/user/Desktop/Data Mining/farmers_market.csv", na = na.vals)

```

## Question 1 Add a new column Season1Days that contains the number of days a market is opened per week (for the dates it is open).
```{r}
#regex to recognize day of week
re_season <- str_c(
  "(?i)(Mon|Tue|Wed|Thu|Fri|Sat|Sun)")

#Add column with counts per week for season 1
fmarkets <- fmarkets %>%
  mutate(
    Season1Days = Season1Time %>%
      str_count(re_season))
```

## Question 2  Add a new column WeekendOpen indicating whether a market opens during weekends in Season1.
```{r}
#use ifelse to change true false to yes no when weekend detected
fmarkets<- fmarkets %>% 
  mutate(WeekendOpen = ifelse(str_detect(Season1Time, "(?i)(Sat|Sun)"), "YES", "NO"))
```

##Question 3 Find out which markets close before 6PM, and which open only for fewer than 4 hours a day. For simplicity, consider only Season1Time. For markets with different open hours across a week, use the average length of open hours for the days they actually open
```{r}
# Pull out column desired to work with outside of the dataframe.
# Make formats the same for time before using strptime
dfTime<-as.data.frame(fmarkets$Season1Time)
colnames(dfTime)<-"MarketTime"
dfTime$MarketTime<-as.character(dfTime$MarketTime)
dfTime$MarketTime<-gsub("(M-)", "M - ", dfTime$MarketTime)
dfTime$MarketTime<-gsub("([a-z]:)", "\\1 ", dfTime$MarketTime)
dfTime$MarketTime<-gsub(";", "; ", dfTime$MarketTime)

#strptime only recognizing one time at a time.  Get all of the times in separate columns
dfTime$First<-strptime(dfTime$MarketTime, "%a: %I:%M %p")
dfTime$second<-strptime(dfTime[,1], "%a: %I:%M %p - %I:%M %p;")
dfTime$third<-strptime(dfTime[,1], 
                       "%a: %I:%M %p - %I:%M %p; %a: %I:%M %p")
dfTime$fourth<-strptime(dfTime[,1], 
                        "%a: %I:%M %p - %I:%M %p; %a: %I:%M %p - %I:%M %p;")
dfTime$fifth<-strptime(dfTime[,1], 
                       "%a: %I:%M %p - %I:%M %p; %a: %I:%M %p - %I:%M %p; %a: %I:%M %p")
dfTime$sixth<-strptime(dfTime[,1], 
                       "%a: %I:%M %p - %I:%M %p; 
%a: %I:%M %p - %I:%M %p; %a: %I:%M %p - %I:%M %p;")
dfTime$seventh<-strptime(dfTime[,1], 
"%a: %I:%M %p - %I:%M %p; 
%a: %I:%M %p - %I:%M %p; %a: %I:%M %p - %I:%M %p; %a: %I:%M %p")
dfTime$eighth<-strptime(dfTime[,1], 
"%a: %I:%M %p - %I:%M %p; 
%a: %I:%M %p - %I:%M %p; %a: %I:%M %p - %I:%M %p; 
%a: %I:%M %p - %I:%M %p;")
dfTime$ninth<-strptime(dfTime[,1],
"%a: %I:%M %p - %I:%M %p; 
%a: %I:%M %p - %I:%M %p; %a: %I:%M %p - %I:%M %p; 
%a: %I:%M %p - %I:%M %p; %a: %I:%M %p")                        
                        
dfTime$tenth<-strptime(dfTime[,1],                        
"%a: %I:%M %p - %I:%M %p; 
%a: %I:%M %p - %I:%M %p; %a: %I:%M %p - %I:%M %p; 
%a: %I:%M %p - %I:%M %p; %a: %I:%M %p - %I:%M %p;")

dfTime$eleventh<-strptime(dfTime[,1],                        
"%a: %I:%M %p - %I:%M %p; 
%a: %I:%M %p - %I:%M %p; %a: %I:%M %p - %I:%M %p; 
%a: %I:%M %p - %I:%M %p; %a: %I:%M %p - %I:%M %p;
%a: %I:%M %p")
dfTime$twelve<-strptime(dfTime[,1],                        
"%a: %I:%M %p - %I:%M %p; 
%a: %I:%M %p - %I:%M %p; %a: %I:%M %p - %I:%M %p; 
%a: %I:%M %p - %I:%M %p; %a: %I:%M %p - %I:%M %p;
%a: %I:%M %p - %I:%M %p;")
dfTime$thirt<-strptime(dfTime[,1],                        
"%a: %I:%M %p - %I:%M %p; 
%a: %I:%M %p - %I:%M %p; %a: %I:%M %p - %I:%M %p; 
%a: %I:%M %p - %I:%M %p; %a: %I:%M %p - %I:%M %p;
%a: %I:%M %p - %I:%M %p; %a: %I:%M %p")
dfTime$fourt<-strptime(dfTime[,1],                        
"%a: %I:%M %p - %I:%M %p; 
%a: %I:%M %p - %I:%M %p; %a: %I:%M %p - %I:%M %p; 
%a: %I:%M %p - %I:%M %p; %a: %I:%M %p - %I:%M %p;
%a: %I:%M %p - %I:%M %p; %a: %I:%M %p - %I:%M %p;")

# calculate how long market is open as numeric
dfTime$DiffOne<- as.numeric(abs(difftime(dfTime$First, dfTime$second, units = "hours"))) 
dfTime$DiffTwo<-as.numeric(abs(difftime(dfTime$third, dfTime$fourth, units = "hours")))
dfTime$DiffThree<-as.numeric(abs(difftime(dfTime$fifth, dfTime$sixth, units = "hours")))
dfTime$DiffFour<-as.numeric(abs(difftime(dfTime$seventh, dfTime$eighth, units = "hours")))
dfTime$DiffFive<-as.numeric(abs(difftime(dfTime$ninth, dfTime$tenth, units = "hours")))
dfTime$DiffSix<-as.numeric(abs(difftime(dfTime$eleventh, dfTime$twelve, units = "hours")))
dfTime$DiffSeven<-as.numeric(abs(difftime(dfTime$thirt, dfTime$fourt, units = "hours")))

#Find average time and ignore column if it contains NA
dfTime <- transform(dfTime, meanOpenTime = rowMeans(dfTime[,16:22], na.rm = TRUE))

#Find if any market is open 4 or less hours
dfTime$OpenLessThanFourHours<-ifelse(dfTime$meanOpenTime < 4, "YES", "NO")

dfTime<-unite(dfTime,
              col = "timesCombined",
              from = c("First", "second", "third", "fourth",    
                        "fifth", "sixth", "seventh", "eighth",                              "ninth", "tenth", "eleventh", "twelve",                             "thirt", "fourt" ),
              sep = ",")

str(dfTime$timesCombined)
condition<-str_detect(dfTime$timesCombined, "18:00|18:30|19:00|19:30|20:00|20:30|21:00|21:30|22:00|22:30|23:00|23:30|24:00")
dfTime$ClosingTime6<-ifelse(condition == "TRUE", "Close at or after 6", "Close before 6")

#Add the columns back into the main df
fmarkets$MeanOpenTime<-dfTime$meanOpenTime
fmarkets$OpenLessThanFourHours<-dfTime$OpenLessThanFourHours
fmarkets$ClosingTime6<-dfTime$ClosingTime6

```


##The seasons are not standardized and would make analysis difficult. Create four new columns for four seasons (Spring, Summer, Fall, Winter), indicating whether a market is available in that season. Also, create two additional columns HalfYear and YearRound to identify those who open across seasons. Define "half year" and "year round" on your own terms, but explain them before you write the code (or as comments in your code). (Hint: you may want to create even more auxiliary columns, Season1BeginDate and Season1EndDate for example.)

```{r}
#copy data frame to work with
df4 <- as.data.frame(fmarkets$Season1Date)

#split on to and separate into different columns
df4<-str_split_fixed(df4$`fmarkets$Season1Date`, "to", 2)

#force into data frame
df4<-as.data.frame(df4)

#insert column names
colnames(df4)<- c("open.date", "close.date")

#use lubridate to format
df4$open.date.clean<-mdy(df4$open.date)
df4$close.date.clean<-mdy(df4$close.date)

#replace NA values which should have dates
df4$open.date.clean <- ifelse(is.na(df4$open.date.clean), df4$open.date, df4$open.date.clean)
df4$close.date.clean <- ifelse(is.na(df4$close.date.clean), df4$close.date, df4$close.date.clean)

str(df4$open.date.clean)

#get rid of values which should not exist 
df4$open.date.clean<-na_if(df4$open.date.clean, 1)
df4$close.date.clean<-na_if(df4$close.date.clean, 1)

#put back in date format
df4$open.date.clean<-as.Date(df4$open.date.clean, origin ="1970-01-01")
df4$close.date.clean<-as.Date(df4$close.date.clean, origin ="1970-01-01")

#Find time differences in months
df4$time.diff.months<-round((df4$close.date.clean - df4$open.date.clean)/(365.25/12))
df4$time.diff.months<-abs(df4$time.diff.months)

#load library zoo 
library(zoo)

df4$ym1 <- as.yearmon(as.character(df4$open.date.clean), "%Y-%m-%d") 
df4$ym2 <- as.yearmon(as.character(df4$close.date.clean), "%Y-%m-%d") 

sequenceFunction <- function (x,y) {
  return(tryCatch(seq(x, y, 1/12), error=function(e) NULL))
}

datesinmonths<-mapply(sequenceFunction, df4$ym1, df4$ym2)

#get list of list as character
df4$DateRangeOpen<-lapply(datesinmonths, as.character)

#unlist the list and split on the comma
df4<-df4 %>%
  mutate(DateRangeOpen = map(DateRangeOpen, toString ) ) %>%
  separate(DateRangeOpen, into = c("M1", "M2", "M3", "M4", "M5", "M6", 
                                   "M7", "M8", "M9", "M10", "M11", "M12"), 
           sep = "(,)", convert = TRUE)

mode(df4$M1)

df4$CombinedDates <- paste(df4$M1, df4$M2, df4$M3, df4$M4, df4$M5, df4$M6, df4$M7, df4$M8,
                      df4$M9, df4$M10, df4$M11, df4$M12, sep=",")
df4$CombinedDates<-gsub("NA,", "", df4$CombinedDates)
df4$CombinedDates<-gsub(",NA", "", df4$CombinedDates)
mode(df4$CombinedDates)
df4$OpenDuringSeasons<-gsub("Dec|Jan|Feb", "Winter", df4$CombinedDates)
df4$OpenDuringSeasons<-gsub("Mar|Apr|May", "Spring", df4$OpenDuringSeasons)
df4$OpenDuringSeasons<-gsub("Jun|Jul|Aug", "Summer", df4$OpenDuringSeasons)
df4$OpenDuringSeasons<-gsub("Sep|Oct|Nov", "Fall", df4$OpenDuringSeasons)
df4$OpenDuringSeasons<-gsub(" \\d+","", df4$OpenDuringSeasons)

#remove duplicates
df4$OpenDuringSeasons <- sapply(df4$OpenDuringSeasons, function(x) paste(unique(unlist(str_split(x,", "))), collapse = ", "))

df4<- df4 %>% 
  mutate(Summer = ifelse(str_detect(OpenDuringSeasons, "(Summer)"), "YES", ""))

df4<- df4 %>%
  mutate(Spring = ifelse(str_detect(OpenDuringSeasons, "(Spring)"), "YES", ""))

df4<- df4 %>% 
  mutate(Winter = ifelse(str_detect(OpenDuringSeasons, "(Winter)"), "YES", ""))

df4<- df4 %>% 
  mutate(Fall = ifelse(str_detect(OpenDuringSeasons, "(Fall)"), "YES", ""))

#if market open for 6 months assign yes to half year
df4<- df4 %>% 
  mutate(HalfYear =ifelse(time.diff.months == 6, "OPEN", "NOT OPEN"))

#if market open more than 10 months assign yes to full year
df4<- df4 %>% 
  mutate(YearRound =ifelse(time.diff.months>10, "OPEN", "NOT OPEN"))

#add new data to existing fmarkets data frame
fmarkets$Winter<-df4$Winter
fmarkets$Spring<-df4$Spring
fmarkets$Summer<-df4$Summer
fmarkets$Fall<-df4$Fall
fmarkets$HalfYear<-df4$HalfYear
fmarkets$YearRound<-df4$YearRound
```

## Question 5 Open question: explore the new variables you just created. Aggregate them at different geographic levels, or some other categorical variable. What can you discover?

```{r}
#explore how how many markets are open less than 4 hours vs more than 4 hours
mode(fmarkets$OpenLessThanFourHours)
length(grep("NO", fmarkets$OpenLessThanFourHours))
length(grep("YES", fmarkets$OpenLessThanFourHours))
```
More markets are open for less than 4 hours a day than open for more than 4 hours a day

```{r}
#explore how many markets close before 6pm vs at or after 6pm
mode(fmarkets$ClosingTime6)
length(grep("Close before 6", fmarkets$ClosingTime6))
length(grep("Close at or after 6", fmarkets$ClosingTime6))
```
More markets close before 6pm than closing at 6pm or after 6pm.

```{r}
#explore which season markets are open the most
length(grep("YES", fmarkets$Fall))
length(grep("YES", fmarkets$Winter))
length(grep("YES", fmarkets$Spring))
length(grep("YES", fmarkets$Summer))
```
Markets are open most often (in order) are Summer, Fall, Spring, Winter.